\LARGE \bf Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation of RGB-D Images

### 2014-01-25 paper.tex
camera ready
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4836 b2511e17-9ac0-488b-bc21-1d6d82e5baa2

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/teaser}
    \end{center}
    \caption{%
        Overview of the proposed semantic segmentation method.\label{teaser}
    \figlabel{height}}
\end{figure}

@@

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor.

@@

All hyper parameters were adjusted using 5-fold cross-validation. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid-search.
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.

@@

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{images/vertical_alignment}
    \includegraphics[width=.45\linewidth]{images/depth_difference}
    \end{center}
    \vspace{-3mm}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See \secref{results} for details.
    \figlabel{weights}}
    \vspace{-3mm}
\end{figure}

\Figref{weights} visualizes some learned potential functions.
Higher values correspond to favored configurations. One can see that the
vertical alignment potential expresses that the floor is much more likely to be below
other classes. It also encodes the fact that props
rest on furniture, but not the other way around.
The potential of the depth feature encodes, for example, that the ground
is usually behind the other classes, while furniture is in front of structures,
such as the wall.
Interestingly, the potential functions are not anti-symmetric, and forcing them to be so
degrades the results. This suggests that the direction of connecting edges, going from
the top left to the bottom right in the image, is also exploited by the potentials.

@@

\section{SUMMARY AND DISCUSSION}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the art on the NYU V2
dataset for semantic annotation of structure classes.

>>>>

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/teaser}
    \end{center}
    \caption{%
        Overview of the proposed semantic segmentation method.\figlabel{teaser}
    }
\end{figure}

@@

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor. The dataset is split into 795 images for training and 654 images for testing.

@@

All hyper parameters were adjusted using 5-fold cross-validation on the training set. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid-search. 
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.
Training of the random forests took about 15 minutes on a NVIDIA GeForce GTX
Titan. Training the structural SVM took 45 minutes on a Xeon X5650 CPU.
Prediction using only the random forest takes 33ms on average, while segmentation and prediction
using the structural SVM approximately take an additional 500ms.

@@

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.
\Figref{weights} visualizes some learned potential functions.

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{images/vertical_alignment}
    \includegraphics[width=.45\linewidth]{images/depth_difference}
    \end{center}
    \vspace{-6mm}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See \secref{results} for details.
    \figlabel{weights}}
    \vspace{-6mm}
\end{figure}

Higher values correspond to favored configurations. One can see that the
vertical alignment potential expresses that the floor is much more likely to be below
other classes. It also encodes the fact that props
rest on furniture, but not the other way around.
The potential of the depth feature encodes, for example, that the ground
is usually behind the other classes, while furniture is in front of structures,
such as the wall.
Interestingly, the potential functions are not anti-symmetric, and forcing them to be so
degrades the results. This suggests that the direction of connecting edges, going from
the top left to the bottom right in the image, is also exploited by the potentials.

@@

\section{SUMMARY AND DISCUSSION}
\enlargethispage{5mm}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model on the output of an
efficient GPU implementation of random forest, and model spatial neighborhood
using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the-art on the NYU V2
dataset for semantic annotation of structure classes.
